---
title: "Stats 487 Activity 3.1"
author: 'Vinh Hang'
date: '2022-10-26'
output:
tufte::tufte_html: default
---

# Stats 487 Activity 3.1

Outline:

1. [Project Data]
2. [Task 1]
3. [Task 2]
4. [References]

R requirements: ggplot2, dplyr

```{r imports, warning = FALSE, echo = FALSE, message = FALSE}
library(tibble)
library(ISLR)
library(MASS)
library(sfsmisc)
library(dplyr)
library(lmtest)

# turn off scientific notation
options(scipen = 999)
```

## Project Data

```{r load-data, warning = FALSE, message = FALSE, echo = FALSE}
# load data
df <- as_tibble(read.csv('D:\\DataspellProjects\\MAS_Fall_22\\Stat_487\\data\\health.csv', stringsAsFactors = T))
df['Female'] <- as.factor(df$Female)
df['FedRegion'] <- as.factor(df$FedRegion)

head(df)
```

## Task 1

#### 1. Uni-variate descriptive statistics

```{r univariate-descriptive-statistic}
# discriptive stats
summary(df[c('TotCostPer', 'Testusers', 'Age')])

# plots 3 in one
par(mfrow = c(1, 3))

# hist plot with distribution on top
hist(df$TotCostPer, freq = FALSE, , main = 'Histogram of TotCostPer', xlab = 'TotCostPer')
lines(density(df$TotCostPer), col = 'red', lwd = 2)

hist(df$Testusers, freq = FALSE, , main = 'Histogram of Testusers', xlab = 'Testusers')
lines(density(df$Testusers), col = 'red', lwd = 2)

hist(df$Age, freq = FALSE, main = 'Histogram of Age', xlab = 'Age')
lines(density(df$Age), col = 'red', lwd = 2)

```

#### 2. Bi-variate descriptive statistics

```{r bivariate-statistics}
# corr matrix
cor(df[c('TotCostPer', 'Testusers', 'Age')])

# pair plot
pairs(df[c('TotCostPer', 'Testusers', 'Age')])
```

#### 3. OLS

```{r ols}
set.seed(27)

# regression
reg_ols <- lm(TotCostPer ~ Testusers + Age, data=df)

# summary
summary(reg_ols)
```

#### 4. Checking assumptions

```{r checking-assumptions}
# 4 in 1 checks
par(mfrow = c(2, 2))
plot(reg_ols)
```

#### 5. Testing assumptions

```{r Shapiro-Wilk-normality-test}
sresid <- studres(reg_ols)
shapiro.test(sresid)
```

```{r ncvTest-Homoscedasticity}
ncvTest(reg_ols)
```

```{r durbin-Watson-auto}
durbinWatsonTest(reg_ols)
```

#### 6. Robust regression

```{r robust-regression}
reg_robust <- rlm(TotCostPer ~ Testusers + Age, data=df)
summary(reg_robust)
```

#### 7. Bi-variate descriptive statistics


```{r test-Testusers}
f.robftest(reg_robust, var = "Testusers")
```

```{r test-Age}
f.robftest(reg_robust, var = "Age")
```

```{r}
sqrt(mean(reg_robust$residuals^2))
```

```{r plot-robust}
par(mfrow=c(2,2))
plot(reg_robust,which=1:3)
```

## Task 2

#### 8. Uni-variate descriptive statistics

```{r univariate-descriptive-statistic-3var}
# discriptive stats
summary(df[c('TotCostPer', 'Female', 'FedRegion')])

# plots 3 in one
par(mfrow = c(1, 3))

# hist plot with distribution on top
hist(df$TotCostPer, freq = FALSE, , main = 'Histogram of TotCostPer', xlab = 'TotCostPer')
lines(density(df$TotCostPer), col = 'red', lwd = 2)

plot(df$Female, main = 'Barplot of Female', xlab = 'Female')

plot(df$FedRegion, main = 'Barplot of FedRegion', xlab = 'FedRegion')

```

#### 9. Bi-variate descriptive statistics


```{r}
# plots 3 in one
par(mfrow = c(1, 2))
# ggboxplot(df, x = "Female", y = "TotCostPer", color = "Female", palette = c("#00AFBB", "#E7B800"))
# ggboxplot(df, x = "FedRegion", y = "TotCostPer", color = "FedRegion", palette = c("#00AFBB", "#E7B800"))
boxplot(df$TotCostPer ~ df$Female, main = 'Boxplot of TotCostPer/Female', ylab = 'TotCostPer', xlab = 'Female', col = c("#00AFBB", "#E7B800"), frame=FALSE)
boxplot(df$TotCostPer ~ df$FedRegion, main = 'Boxplot of TotCostPer/FedRegion', ylab = 'TotCostPer', xlab = 'FedRegion',frame=FALSE)
```


#### 10. ANOVA

```{r anova}
res_aov <- aov(TotCostPer ~ Female + FedRegion, data = df)
summary(res_aov)
```

#### 11. Check assumptions


```{r}
leveneTest(TotCostPer ~ Female * FedRegion, data = df)
```

```{r Shapiro-Wilk-normality-test}
sresid <- residuals(res_aov)
shapiro.test(sresid)
```

```{r}
par(mfrow=c(2,2))
plot(res_aov)
```

```{r anova-interactions-2}
res_aov_int <- aov(TotCostPer ~ Female + FedRegion + Female:FedRegion, data = df)
summary(res_aov_int)
```

#### 12. Checking Assumptions
#### 13. Robust
#### 14. Bi-variate descriptive statistics

```{r}
res_aov <- aov(TotCostPer ~ Female + FedRegion, data = df)
Anova(res_aov, type='III')
```

```{r}
res_aov <- aov(TotCostPer ~ Female, data = df)
Anova(res_aov, type='III')
```


```{r}
par(mfrow=c(2,2))
plot(res_aov)
```

## References

For MANOVA, LDA, K-mean algorithm, please refer to the links below

[PSU Stats 508](https://online.stat.psu.edu/stat508/)

An Introduction to Statistical Learning for more in-dept reference

[ISLR](https://www.statlearning.com/)

[LDA example](https://pages.cms.hu-berlin.de/EOL/gcg_quantitative-methods/Lab11_LDA_Model-assessment.html)

[MANOVA exmample](https://www.datanovia.com/en/lessons/one-way-manova-in-r/#assumptions-and-preleminary-tests)

[K-mean exmample](https://www.datanovia.com/en/lessons/k-means-clustering-in-r-algorith-and-practical-examples/)


