---
title: "Stats 580 Second Project"
author: 'Vinh Hang'
date: '2022-10-08'
output:
tufte::tufte_html: default
---

# Stat 580 Project 2

## Executive Summary



Outline:

1. [Project Description]
2. [Research Questions and Statistical Approaches]
3. [Variables]
4. [EDA]
5. [Statistical Analysis]
6. [Recommendations]
7. [Resources]
8. [Considerations]

R requirements: psych, ggplot2, dplyr, gridExtra, grid, ggpubr, patchwork, table, qacr, factoextra, klaR, psych, MASS, devtools, rstatix, ggbiplot, caret

```{r imports, warning = FALSE, echo = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(norm)
library(Hmisc)
```

## Project Description


## Research Questions and Statistical Approaches


## Variables

```{r load-data, warning = FALSE, message = FALSE, echo = FALSE}
# add neighborgood indicator and KitchenAbvGr for CollegeCr dataset
df_college <- read.csv('D:\\DataspellProjects\\MAS_Fall_22\\Stat_580\\project_2\\data\\CollegeCr.csv', stringsAsFactors = T)
df_college['neighborhood'] <- 'CollegeCr'
df_college['KitchenAbvGr'] <- NA

# add neighborgood indicator and BsmtUnfSF for Edwards dataset
df_edwards <- read.csv('D:\\DataspellProjects\\MAS_Fall_22\\Stat_580\\project_2\\data\\Edwards.csv', stringsAsFactors = T)
df_edwards['neighborhood'] <- 'Edwards'
df_edwards['BsmtUnfSF'] <- NA

# add neighborgood indicator for OldTown dataset
df_oldtowm <- read.csv('D:\\DataspellProjects\\MAS_Fall_22\\Stat_580\\project_2\\data\\OldTown.csv', stringsAsFactors = T)
df_oldtowm['neighborhood'] <- 'OldTown'

# scale data
df_full <- rbind(df_college, df_edwards, df_oldtowm)
head(df_full)
```


```{r}
# Split Lot Info into LotConfig, LotShape, LotArea, LotFrontage
df_full[c('LotConfig', 'LotShape', 'LotArea', 'LotFrontage')] <- str_split_fixed(df_full$LotInfo, ';', 4)
df_full['LotArea'] <- as.numeric(unlist(df_full['LotArea']))
df_full['LotFrontage'] <- as.numeric(unlist(df_full['LotFrontage']))
df_full['LotConfig'] <- as.character(unlist(df_full['LotConfig']))
df_full['LotShape'] <- as.character(unlist(df_full['LotShape']))

# Split Exterior Info into Exterior1st, ExterQual, ExterCond
df_full[c('Exterior1st', 'ExterQual', 'ExterCond')] <- str_split_fixed(df_full$Exterior, ';', 3)
df_full['Exterior1st'] <- as.character(unlist(df_full['Exterior1st']))
df_full['ExterQual'] <- as.character(unlist(df_full['ExterQual']))
df_full['ExterCond'] <- as.character(unlist(df_full['ExterCond']))

# Drop redundant
df_full <- subset(df_full, select = -c(LotInfo, Exterior))

# convert to categorical
head(df_full)
```


```{r}
# impute missing values numerica with median
df_full$LotFrontage <- impute(df_full$LotFrontage, median)
df_full$BsmtUnfSF <- impute(df_full$BsmtUnfSF, median)
df_full$KitchenAbvGr <- impute(df_full$KitchenAbvGr, median)

# impute manually categorical
df_full$BsmtCond[is.na(df_full$BsmtCond)] <- 'TA'

sort(colSums(is.na(df_full)), decreasing = TRUE)
```

```{r}
# save full train dataset
write.csv(df_full, "D:\\DataspellProjects\\MAS_Fall_22\\Stat_580\\project_2\\data\\house_training.csv")
```

## EDA



## Statistical Analysis

